---
title: "PHS_project_02"
author: "Jonny Nelson"
date: "02/01/2022"
output: html_document
---

```{r}
library(janitor)
library(tidyverse)
library(lubridate)
library(tsibbledata)
library(tsibble)
library(fable)
library(slider)
library(feasts)
library(tidyquant)
library(sf)
library(rgdal)

ae <- read_csv("monthly_ae_waitingtimes_202110.csv")

nhs_boards <- read_csv("PHS_a_and_e_health_board_keys.csv")

ae_clean <- ae %>%
  clean_names() %>%
  mutate(date = ym(month),
         year_month = as.yearmon(date),
         year = year(date),
         month = month(date))


glimpse(ae)

unique(ae$Month)

# Month is YearMonth format (dbl)

unique(ae$Country)

# Scotland

# TreakmentLocation == NHS Hospital

unique(ae$DepartmentType) 

# "Minor Injury Unit or Other" "Emergency Department" 

# Number of Attendances

head(ae, 100)

# NumberOfAttendancesAggregate split by health board, department type

# AttendanceGreater8/12 Hours - number

unique(ae_clean$number_of_attendances_aggregate)
```

# Histogram of Admissions

```{r}
ae_clean %>%
  select(date, number_of_attendances_aggregate, department_type) %>%
  filter(date >= "2017-07-01") %>%
  group_split(date) %>%
  # mutate(
  #   moving_avg = slide_dbl(
  #     .x = number_of_attendances_aggregate, 
  #     .f = ~ mean(., na.rm = TRUE),
  #     .before = 100,
  #     .after = 100
  #   )
  # ) %>%
  ungroup()%>%
  ggplot(aes(x = date, y = number_of_attendances_aggregate)) +
  geom_col(aes(x = date, y =  number_of_attendances_aggregate, fill = department_type)) +
  geom_ma(ma_fun = SMA, n = 30, colour = "red") +
  ggtitle("Time Series for Number of Attendances") +
  labs(x = "Date", y = "Number of Attendances")

# Need to get the moving average to work
```

# Getting the moving average to work

```{r}
ae_clean %>%
  select(date, number_of_attendances_aggregate, department_type) %>%
  filter(date >= "2017-07-01") %>%
  mutate(
    moving_avg = slide_dbl(
      .x = number_of_attendances_aggregate, 
      .f = ~ mean(., na.rm = TRUE),
      .before = 100,
      .after = 100
    )
  ) %>%
  ggplot() +
  geom_col(aes(x = date, y =  number_of_attendances_aggregate, fill = department_type)) +
  ggtitle("Time Series for Number of Attendances") +
  labs(x = "Date", y = "Number of Attendances")

# Need to get the moving average to work
```
# Transforming data into per year basis

```{r}
ae_clean %>%
  select(date, number_of_attendances_aggregate, department_type, hbt) %>%
  filter(date >= "2017-07-01") %>%
  mutate(
    moving_avg = slide_dbl(
      .x = number_of_attendances_aggregate, 
      .f = ~ mean(., na.rm = TRUE),
      .before = 100,
      .after = 100
    )
  ) %>%
  ggplot() +
  geom_col(aes(x = date, y =  number_of_attendances_aggregate, fill = department_type)) +
  ggtitle("Time Series for Number of Attendances") +
  labs(x = "Date", y = "Number of Attendances")
```

# Number of Attendances by HBT

```{r}
ae_clean %>%
  select(date, number_of_attendances_aggregate, department_type, hbt) %>%
  filter(date >= "2017-07-01") %>%
  mutate(
    moving_avg = slide_dbl(
      .x = number_of_attendances_aggregate, 
      .f = ~ mean(., na.rm = TRUE),
      .before = 100,
      .after = 100
    )
  ) %>%
  ggplot() +
  geom_col(aes(x = date, y =  number_of_attendances_aggregate, fill = hbt)) +
  ggtitle("Time Series for Number of Attendances") +
  labs(x = "Date", y = "Number of Attendances")
```

# Attendances over 8 Hours

```{r}
ae_clean %>%
  select(date, attendance_greater8hrs, department_type, hbt) %>%
  filter(date >= "2017-07-01") %>%
  ggplot() +
  geom_col(aes(x = date, y =  attendance_greater8hrs, fill = department_type)) +
  ggtitle("Time Series for Number of Attendances over 8 Hours") +
  labs(x = "Date", y = "Number of Attendances over 8 Hours")
```


# Time Series for Number of Attendances over 12 Hours

```{r}
ae_clean %>%
  select(date, attendance_greater12hrs, department_type, hbt) %>%
  # filter(date >= "2017-07-01") %>%
  ggplot() +
  geom_col(aes(x = date, y =  attendance_greater12hrs, fill = department_type)) +
  ggtitle("Time Series for Number of Attendances over 12 Hours") +
  labs(x = "Date", y = "Number of Attendances over 12 Hours")
```

```{r}
# Need to convert the ae_clean into a tsibble

ae_tsibble <- ae_clean %>%
  tsibble(key = NULL, index = date)

# Cannot make into a tsibble as there are duplicate entries for time and cannot collapse the data.
```

# Cannot make into a tsibble

# Attendance Over 12 Hours per Year

```{r}

clean_ae %>%
  filter(year_month >= "Jan 2016" & department_type == "Emergency Department") %>%
  group_by(year_month) %>%
  summarise(num_attendance = sum(attendance_greater12hrs, na.rm = TRUE)) %>%
  ggplot(aes(x = month(year_month,
                       label = TRUE,
                       abbr = TRUE),
             y = num_attendance,
         group = factor(year(year_month)),
         colour = factor(year(year_month))
         )) +
  geom_line() +
  geom_point() +
  labs(x = "Month",
       y = "Attendance Over 12 Hours",
       colour = "Year") +
  theme_classic()

# unique(ae_clean$department_type)

  #   mutate(
  #   moving_avg = slide_dbl(
  #     .x = number_of_attendances_aggregate, 
  #     .f = ~ mean(., na.rm = TRUE),
  #     .before = 100,
  #     .after = 100
  #   )
  # ) %>%

```

# Attendance Over 8 Hours per Year

```{r}
ae_clean %>%
  filter(year_month >= "Jan 2016") %>%
  group_by(year_month) %>%
  summarise(num_attendance = sum(attendance_greater8hrs, na.rm = TRUE)) %>%
  ggplot(aes(x = month(year_month,
                       label = TRUE,
                       abbr = TRUE),
             y = num_attendance,
         group = factor(year(year_month)),
         colour = factor(year(year_month))
         )) +
  geom_line() +
  geom_point() +
  labs(x = "Month",
       y = "Attendance Over 8 Hours",
       colour = "Year") +
  theme_classic()

```

# Wrangling Ricardos Work

# 2021 outlier year for Summer Attendances over 12 Hours

```{r}
      ae_clean %>%
        group_by(month, year) %>%
        summarise(attendance = sum(attendance_greater12hrs, na.rm = T)) %>%
        mutate(winter_flag = case_when(month %in% c(1, 2, 3, 10, 11, 12) ~ "Yes",
                                       TRUE ~ "No")) %>%
        group_by(winter_flag, year) %>%
        summarise(av_attendance = mean(attendance)) %>%
        arrange(year) %>%
        filter(winter_flag == "No") %>%
        ggplot() +
        aes(x = year, y = av_attendance) +
        geom_col() +
        labs(x = "Year",
             y = "Attendance Over the Previous Year") +
        ggtitle("Number of Attendances compared to Previous Year") +
        theme_classic() +
        theme(plot.title = element_text(size = 16, hjust = 0.5))
```

# Winter Attendances over 12 Hours 

```{r}
      ae_clean %>%
        group_by(month, year) %>%
        summarise(attendance = sum(attendance_greater12hrs, na.rm = T)) %>%
        mutate(winter_flag = case_when(month %in% c(1, 2, 3, 10, 11, 12) ~ "Yes",
                                       TRUE ~ "No")) %>%
        group_by(winter_flag, year) %>%
        summarise(av_attendance = mean(attendance)) %>%
        arrange(year) %>%
        filter(winter_flag == "Yes") %>%
        ggplot() +
        aes(x = year, y = av_attendance) +
        geom_col() +
        labs(x = "Year",
             y = "Attendance Over the Previous Year") +
        ggtitle("Number of Attendances compared to Previous Year") +
        theme_classic() +
        theme(plot.title = element_text(size = 16, hjust = 0.5))
```

# Using Winter Flag as a fill to compare

```{r}
      ae_clean %>%
        group_by(month, year) %>%
        summarise(attendance = sum(attendance_greater12hrs, na.rm = T)) %>%
        mutate(winter_flag = case_when(month %in% c(1, 2, 3, 10, 11, 12) ~ "Yes",
                                       TRUE ~ "No")) %>%
        group_by(winter_flag, year) %>%
        summarise(av_attendance = mean(attendance)) %>%
        arrange(year) %>%
        ggplot(aes(x = year,
                   y = av_attendance,
                   fill = winter_flag),
               position = dodge) +
        geom_col() +
        labs(x = "Year",
             y = "Attendance Over the Previous Year") +
        ggtitle("Number of Attendances compared to Previous Year") +
        theme_classic() +
        theme(plot.title = element_text(size = 16, hjust = 0.5))
```

```{r}
      ae_clean %>%
        filter(year < 2021) %>%
        group_by(month, year) %>%
        summarise(attendance = sum(attendance_greater12hrs, na.rm = T)) %>%
        mutate(winter_flag = case_when(month %in% c(1, 2, 3, 10, 11, 12) ~ "Yes",
                                       TRUE ~ "No")) %>%
        group_by(winter_flag, year) %>%
        summarise(av_attendance = mean(attendance)) %>%
        arrange(year) %>%
        mutate(percent_change = ((av_attendance/lag(av_attendance)-1)*100)) %>%
        filter(winter_flag == "Yes") %>%
        ggplot(aes(x = year,
                   y = percent_change),
               position = dodge) +
        geom_col(fill = "steel blue",
           colour = "black") +
        labs(x = "Year",
             y = "Attendance Over the Previous Year") +
        ggtitle("Number of Attendances compared to Previous Year") +
        theme_classic() +
        theme(plot.title = element_text(size = 16, hjust = 0.5))
```

#  Trial to make negative values red

```{r}

# New Object to Hold the Statistical Data

      ae_stats <- ae_clean %>%
        filter(year < 2021) %>%
        group_by(month, year) %>%
        summarise(attendance = sum(attendance_greater12hrs, na.rm = T)) %>%
        mutate(winter_flag = case_when(month %in% c(1, 2, 3, 10, 11, 12) ~ "Yes",
                                       TRUE ~ "No")) %>%
        group_by(winter_flag, year) %>%
        summarise(av_attendance = mean(attendance)) %>%
        arrange(year) %>%
        mutate(percent_change = ((av_attendance/lag(av_attendance)-1)*100),
               sign_percent_change = sign(percent_change))
      ae_stats
```
# Percent Change YoY

```{r}
      ae_stats %>%
        filter(winter_flag == "Yes") %>%
        ggplot(aes(x = year,
                   y = percent_change)) +
        geom_col(aes(fill = interaction(sign_percent_change > 0)), colour = "black") +
        scale_fill_manual(values = c("steel blue", "#FF0000"),
                          breaks = c("TRUE", "FALSE")) +
        scale_x_continuous(breaks = 2008:2020) +
        scale_y_continuous(breaks = c(-50, 0, 50, 100, 150, 200)) +
        labs(x = "Year",
             y = "Percentage Change YoY") +
        ggtitle("A&E No. of Attendances Change YoY") +
        theme_classic() +
        theme(plot.title = element_text(size = 16, hjust = 0.5),
              legend.position = "none")
```


# Summary Statistics Table

```{r}
# Cleaned up table to check

ae_clean %>%
  filter(year < 2021) %>%
  group_by(month, year) %>%
  summarise(attendance = sum(attendance_greater12hrs, na.rm = T)) %>%
  mutate(winter_flag = case_when(month %in% c(1, 2, 3, 10, 11, 12) ~ "Yes",
                                 TRUE ~ "No")) %>%
  group_by(winter_flag, year) %>%
  summarise(av_attendance = mean(attendance)) %>%
  arrange(winter_flag, year) %>%
  mutate(percent_change = ((av_attendance/lag(av_attendance)-1)*100)) %>%
  summarise(total_percent_change = sum(percent_change, na.rm = TRUE)/ 13) %>%
  ggplot() +
  geom_col(aes(x = winter_flag,
               y = total_percent_change),
           fill = "steel blue",
           colour = "black") +
  labs(x = "Winter Flag",
       y = "Average Percent Change Year over Year") +
  ggtitle(" Average Percent Change in A&E Attendance Year over Year")

# Vary Winter Flag to Find Winter and Summer
```



## To Do List

### - Clean the data fully (Date)
### - Wrangle the data to spot trends and patterns
### - Plot Spatial Map
### - z qf

```{r}

```











# Spatial Map

```{r}
library(sf)
library(rgdal)
library(raster)
library(ggplot2)
library(rgeos)
library(mapview)
library(leaflet)
library(broom) # if you plot with ggplot and need to turn sp data into dataframes
options(stringsAsFactors = FALSE)

# Code Chunk did not work

# healthboard_shapefile <- readOGR(dsn = "NHS_healthboard_shapefiles_simple/",
#                                  layer = "NHS_HealthBoards_2019",
#                                  GDAL1_integer64_policy = TRUE)

shapes <- read_sf(dsn = "NHS_healthboard_shapefiles_simple/SG_NHS_HealthBoards_2019", layer = "SG_NHS_HealthBoards_2019")

```


